---
title: "Filters and Joins"
author: "Herb"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

library(tidyverse)
library(readxl)
library(here)
library(kableExtra)
```

Get some fish.csv and kelp_fronds.xlsx

```{r}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_xlsx(here("data", "kelp_fronds.xlsx"), sheet = "abur")
```


### Filter by exact match

Exact match: `==`
```{r}

fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
```

Activity
```{r}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
```

### Filter by multiple criteria

Use `|` (vertical Line) as "or"

```{r}
fish_2sp <- fish %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith")
```

Filtering for multiple matches in same variable: use `%in%`
```{r}
fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

## Activity
create a suset fish_gar_2016

```{r}
fish_gar_2016 <- fish %>% 
  filter(common_name == "garibaldi" & year == 2016)
```

## Filter for "and: conditions
- separate by comma in `filter()`
- use `&`
- use multiple `filter()` expressions

## Exclude filters
use `!=`

### `stringr::str_detect()`
```{r}
fish_black <- fish %>% 
  filter(str_detect(common_name, "black"))
```

### Activity: common_name must include "it"
```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, "it"))
```

### `dplyr::*_join"

full_join

```{r}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))
  
```

### left_join
```{r}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year", "site"))
```

### inner_join
```{r}
kelp_fish_inner <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
```

### join and mutate - pipe togeher multiple steps

- create an object `my_fish_join`

```{r}
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count / total_fronds)
```

## HTML Table using kable + kableExtra

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```
```{r}
ggplot(data = my_fish_join, 
       aes(common_name, fish_per_frond)) +
  geom_col(aes(fill = common_name))
```

