---
title: "Pivot Lobsters"
author: "Herb"
date: "1/27/2020"
output: html_document
---

Attached packages and reading data

```{r setup}
library(tidyverse)
library(readxl)
library(here)
library(skimr) #install.packages("skimr")
library(kableExtra) # install.packages("kableExtra")

# read lobster data
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"), skip = 4)
```

## Viewing data with skimr
```{r}
skimr::skim(lobsters)
```

## Pivot Tables
in R, Pivot tables are `group_by() %>% summarize()`
`n()` function will count observations 

```{r}
lobsters %>% 
  group_by(site, year) %>% 
  summarise(count_by_site_year = n())
```

## Summarize by multiple variables
```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarise(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

## Table formatting with `kableExtra::kable()`
```{r}
siteyear_summary %>%
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.

## Activity
1. build for this analysis and calculate the median lobster size for each site and year (hint: `size_mm` and `median()`
2. make a simple ggplot
3. sync to github

```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarise(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

siteyear_summary %>% 
  ggplot(aes(x = year, y = median_size_mm, color = site))+
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))
```

