---
title: "Creating graphs wth ggplot2"
author: "Herb"
date: "1/27/2020"
output: html_document
---
# Plots with ggplot2

We are going to make plots in R using ghe ggplot2 package from the tidyverse

Install packages:   
- tidyverse: wranling & viz (ggplot)  
- readxl: read excel files  
- here: help with file paths  


```{r setup, message=FALSE}
library(tidyverse) # install.packages("tidyverse")
library(readxl) #install.packages("readxl")
library(here) #install.packages("here")
```

# California National Parks

```{r read-data}
ca_np <- read_csv(here("data", "ca_np.csv"))
c1_np <- read_xlsx(here("data", "ci_np.xlsx"))
ci_np <- read_xlsx(here("data", "ci_np.xlsx"))
```

### Make my own first graph

Line plot year vs visitation in ggplot

```{r}
ggplot(ci_np, aes(year, visitors))+
  geom_point()
```


Make a base layer for plot
```{r}
gg_base <- ggplot(ci_np, aes(year, visitors))

gg_base + geom_point()
```

### Let's start customizing
- line of point color
- line type
- point style: pch
- transparency: aplha

```{r}
gg_base +
  geom_point(
    color = "red",
    pch = 2,
    size = 2)
```
### Update aes based on variables

```{r}
gg_base+
  geom_point(
    aes(size = visitors, color = year))
```

### Use gg-themes for overall updates
- theme_light()
- theme_minimal()
- theme_bw()

```{r}
gg_base+
  geom_area() +
  theme_minimal()
```

### Create a multi-series line plot using data foo all CA parks

```{r}
ggplot(ca_np, aes(year, visitors,
                  color = park_name)) +
  geom_line()
```

```{r}
my_np_graph <- ggplot(ca_np, aes(year, visitors)) +
  geom_point(aes(color = park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free") +
  theme_minimal()

```

## Export my graph as png

```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```

