---
title: "Tidying Session or Something"
author: "Herb"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages
library(tidyverse)
library(janitor) #install.packages("janitor")
library(here)
library(readxl)
```

### Read inverts

```{r}
inverts <- read_xlsx(here("data", "inverts.xlsx"))
```

## `pivot_longer()`

Use single : to indicate a series of cols
```{r}
inverts_long <- pivot_longer(data = inverts,
                             cols = '2016':'2018',
                             names_to = "year",
                             values_to = "sp_count")
```

# back to wide format
```{r}
inverts_wide <- inverts_long %>%
  pivot_wider(names_from = common_name,
              values_from = sp_count)
```

### use `janitor::clean_names()`
```{r}
inverts_wide <- inverts_wide %>%
  clean_names()
```

### use `tidyr::unite()` to combine infornmation from columns
Example: merge site name and year into "site_year"
```{r}
inverts_unite <- inverts_long %>%
  unite(col = "site_year",
        c(site, year),
        sep = "_")
```

### Activity:
create a new object "inverts_moyr" starting from
inverts_long that unites month and year into a single column "mo_yr" sep = "/"
```{r}
invert_moyr <- inverts_long %>% 
  unite(col = "mo_yr",
    c(month, year),
        sep = ".")
?unite
```

### `tidyr::separate()`

```{r}
inverts_sep <- inverts_unite %>% 
  separate(col = "site_year",
           into = c("my_site", "my_year"))
```

###  `stringr` + `mutate()`
use "trees" built-in data set - add 5 to all Heights

```{r}
trees_new <- 
  trees %>% 
  mutate(height_update = Height + 5)
```

### use `str_replace()`

```{r}
ca_abbr <- inverts %>% 
  mutate(common_name = str_replace(common_name, pattern = "california", replacement = "CA"))
```

###
IN THE COMMON_NAME COLUMN, REPLACE "SPINY" with "awesome"
```{r}
awesome_lobsters <- inverts %>% 
  mutate(common_name = str_replace(common_name, pattern = "spiny", replacement = "awesome"))
```

## QUESTION

```{r}
home_data <- c("room", "")

name %>% str_detect("\\(")

name %>% str_locate("\\(")

name_sep <- name %>% 
  separate(name, "\\(")
```

